### 2022-09-05 latex/Appendices/Examples.tex
update

\chapter{Code Examples} % Main chapter title

@@

In this chapter, I will go through some examples, illustrating how to use the code base to learn and solve various puzzles. For each section, simple examples of code will be indicated in \afblue \textbf{python code} \black paragraphs, and can easily be run from command line or copied into a script and run from your favourite IDE.

%-----------------------------------
%	SECTION 1
%-----------------------------------
\section{Puzzles}
Let me start by showing how to construct puzzles, using the Puzzle factory. Notice that in order to run a learner or solver of any kind (assuming of course that they are meant to work on the puzzle type in question), we can just use the exact same code, but just specify \textit{puzzle\_type}  and the expected parameters to construct a puzzle.
\\
\\
For instance, let us create a (n=5, m=6) SP, shuffle it a number of times, and print it:
\afblue
\paragraph{}{\textbf{python code -- sliding puzzle construction}}
\begin{python}
####################################################################
from rubiks.puzzle.puzzle import Puzzle
####################################################################
puzzle_type=Puzzle.sliding_puzzle
n=5
m=6
nb_moves=1000
print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
####################################################################
\end{python}
\black
The output from the above code snippet will look like (subject to randomness):

\begin{figure}[H]
\centering
\includegraphics[scale=0.8]{./Figures/examplespconstruction}
%\decoRule
\caption[Examples]{sliding puzzle construction example}
\label{fig:examplespconstruction}
\end{figure}
.. similarly to construct a (n=2) RC and shuffle it perfectly:
\afblue
\paragraph{}{\textbf{python code -- rubiks cube construction}}
\begin{python}
####################################################################
from math import inf
from rubiks.puzzle.puzzle import Puzzle
####################################################################
puzzle_type=Puzzle.rubiks_cube
n=2
""" Here we use perfect shuffle by spcifying infinite number of shuffles """
nb_moves=inf
print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
####################################################################
\end{python}
\black

@@

.. finally let me construct a (n=3) RC with non-standard-goal:
\afblue
\paragraph{}{\textbf{python code -- non-standard-goal rubiks cube construction}}
\begin{python}
####################################################################
from math import inf
from rubiks.puzzle.puzzle import Puzzle
####################################################################
puzzle_type=Puzzle.watkins_cube
n=2
""" Here we use perfect shuffle by spcifying infinite number of shuffles """
nb_moves=inf
print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
####################################################################
\end{python}
\black

\begin{figure}[H]
\centering
\includegraphics[scale=0.8]{./Figures/examplewcconstruction}
%\decoRule
\caption[Examples]{non-standard-goal rubiks cube construction example}
\label{fig:examplewcconstruction}
\end{figure}

@@

The perfect learner has been discussed in details in \ref{sec:codelearners}. We simply show here how to run it to learn the value function for the (n=2, m=3) SP, that is, to solve all 360 possible configurations via an optimal solver (A{*} with Manhattan heuristic).

\afblue
\paragraph{}{\textbf{python code -- perfect learner}}
\begin{python}
####################################################################
from rubiks.heuristics.heuristic import Heuristic
from rubiks.puzzle.puzzle import Puzzle
from rubiks.learners.learner import Learner
from rubiks.utils.utils import get_model_file_name
####################################################################
action_type=Learner.do_learn
n=2
m=3
puzzle_type=Puzzle.sliding_puzzle
learner_type=Learner.perfect_learner
heuristic_type=Heuristic.manhattan
nb_cpus=4
learning_file_name=get_model_file_name(puzzle_type=puzzle_type,
                                       dimension=(n, m),
                                       model_name=Learner.perfect)
if __name__ == '__main__':
    # we fully solve the 2 x 3 SP ... should take ~5s
    Learner.factory(**globals()).action()
####################################################################
action_type=Learner.do_plot
if __name__ == '__main__':
    # we display the results
    Learner.factory(**globals()).action()
####################################################################
\end{python}
\black
The above snippet of code will solve the 2 by 3 SP and then display the results, showing the distribution of optimal costs, as well as the most difficult puzzle.

\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{./Figures/exampleperfectlearner}
%\decoRule
\caption[Examples]{perfect learner example}
\label{fig:exampleperfectlearner}
\end{figure}

%-----------------------------------
%	SUBSECTION 2.2
%-----------------------------------
\subsection{Deep Learner}
\label{DLSS}
The DeepLearner, also discussed in details in \ref{sec:codelearners}, needs some training data. It could in principle of course be trained on any data, not necessarily optimal data (i.e. generated by an optimal solver). Here however, I make use of the TrainingData class to generate 100 sequences of fully solved (n=5, m=2) SP via A{*} with Manhattan++.

>>>>

\chapter{Examples} % Main chapter title

@@

In this appendix, I will go through many examples, illustrating how to use the code base to create, learn and solve various puzzles. For each section, a snippet of code will be indicated in \afblue \textbf{python code} \black paragraphs, and can easily be run from command line or copied into a script and run from your favourite Python IDE.

%-----------------------------------
%	SECTION 1
%-----------------------------------
\section{Puzzles}
Let me start by showing how to construct puzzles, using the Puzzle factory. Notice that in order to run a learner or solver of any kind (assuming of course that they are meant to work on the puzzle type in question), we can just use the exact same code, simply specifying \textit{puzzle\_type} and the expected parameters to construct a puzzle. The factories will just pick up the relevant parameters indicating the puzzle type and dimension and everything should work seamlessly.
\\
\\
For instance, let us create a 5x6 \textbf{SP}, shuffle it a thousand times, and print it:
\afblue
\paragraph{}{\textbf{python code -- sliding puzzle construction}}
\begin{python}
####################################################################
from rubiks.puzzle.puzzle import Puzzle
####################################################################
puzzle_type=Puzzle.sliding_puzzle
n=5
m=6
nb_moves=1000
print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
####################################################################
\end{python}
\black
The output from the above code snippet will look like (subject to randomness):

\begin{figure}[H]
\centering
\includegraphics[scale=0.8]{./Figures/examplespconstruction}
%\decoRule
\caption[Examples]{scrambled 5x6 \textbf{SP} example}
\label{fig:examplespconstruction}
\end{figure}
.. similarly to construct a perfectly scrambled 2x2x2 \textbf{RC}:
\afblue
\paragraph{}{\textbf{python code -- Rubik's cube construction}}
\begin{python}
####################################################################
from math import inf
from rubiks.puzzle.puzzle import Puzzle
####################################################################
puzzle_type=Puzzle.rubiks_cube
n=2
""" Here we use perfect shuffle by spcifying infinite number of shuffles """
nb_moves=inf
print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
####################################################################
\end{python}
\black

@@

@@

The perfect learner has been discussed in details in \ref{sec:codelearners}. We simply show here how to run it to learn the value function for the 2x3 \textbf{SP}, that is, to solve all 360 possible configurations via an optimal solver (A{*} with Manhattan heuristic).

\afblue
\paragraph{}{\textbf{python code -- perfect learner}}
\begin{python}
####################################################################
from rubiks.heuristics.heuristic import Heuristic
from rubiks.puzzle.puzzle import Puzzle
from rubiks.learners.learner import Learner
from rubiks.utils.utils import get_model_file_name
####################################################################
action_type=Learner.do_learn
n=2
m=3
puzzle_type=Puzzle.sliding_puzzle
learner_type=Learner.perfect_learner
heuristic_type=Heuristic.manhattan
nb_cpus=4
learning_file_name=get_model_file_name(puzzle_type=puzzle_type,
                                       dimension=(n, m),
                                       model_name=Learner.perfect)
if __name__ == '__main__':
    # we fully solve the 2 x 3 SP ... should take ~5s
    Learner.factory(**globals()).action()
####################################################################
action_type=Learner.do_plot
if __name__ == '__main__':
    # we display the results
    Learner.factory(**globals()).action()
####################################################################
\end{python}
\black
The above snippet of code solves the 2x3 \textbf{SP} and then displays the results, showing the distribution of optimal costs, as well as the most difficult puzzle:

\begin{figure}[H]
\centering
\includegraphics[scale=0.41]{./Figures/exampleperfectlearner}
%\decoRule
\caption[Examples]{perfect learner example}
\label{fig:exampleperfectlearner}
\end{figure}

%-----------------------------------
%	SUBSECTION 2.2
%-----------------------------------
\subsection{Deep Learner}
\label{DLSS}
The DeepLearner, also discussed in details in \ref{sec:codelearners}, needs some training data. It could in principle of course be trained on any data, not necessarily optimal data (i.e. generated by an optimal solver). Here however, I make use of the TrainingData class to generate 100 sequences of fully solved 5x2 \textbf{SP} via A{*} with Manhattan++.

### 2022-09-05 latex/Appendices/Kociemba.tex
update

\afblue
\paragraph{}{\textbf{Kociemba bug -- 2x2x2 example of good interface design}}
\begin{python}
#####################################################################
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
from rubiks.solvers.kociembasolver import KociembaSolver
from_kociemba = KociembaSolver.from_kociemba
to_kociemba = KociembaSolver.to_kociemba
#####################################################################
puzzle_type = Puzzle.rubiks_cube
n=2
init_from_random_goal=False
cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
solver_type = Solver.kociemba
solver = Solver.factory(**globals())
print(solver.solve(cube))
#####################################################################
\end{python}
\black

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{./Figures/KociembaBug222}
%\decoRule
\caption[Kociemba Bug]{2x2x2}
\label{fig:KociembaBug222}
\end{figure}
As we can see, hkociemba happily returned a 1-move solution to green F, white U, instead of the costlier solution it would have taken to get red F white U (clearly at a cost of 3 with solution RUD').
\\
\\
In the 3x3x3 case, things are not quite that simple. There is no way to reproduce full cube rotation from faces rotations alone, since the center cubies will never move via the latter, but do via the former. If we run the equivalent of the above problem with the (3x3x3) kociemba library, we silently get to a false solution as is apparent from printing the resulting cube.
\afblue
\paragraph{}{\textbf{Kociemba bug -- 3x3x3 example of \textbf{bad} interface design}}
\begin{python}
#####################################################################
from kociemba import solve
#####################################################################
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.kociembasolver import KociembaSolver
from_kociemba = KociembaSolver.from_kociemba
to_kociemba = KociembaSolver.to_kociemba
#####################################################################
puzzle_type = Puzzle.rubiks_cube
n=3
init_from_random_goal=False
cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
print(cube)
cube_string = to_kociemba(cube)
solution = solve(cube_string)
print('Kociemba Solution:', solution)
print(cube.apply_moves(from_kociemba(solution)))
###################################################################
\end{python}
\black

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{./Figures/KociembaBug333}
%\decoRule
\caption[Kociemba Bug]{3x3x3 Buggy}
\label{fig:KociembaBug333}
\end{figure}
In order to counteract that, and make the comparisons between Kociemba 3x3x3 and other solvers and methods fairer, I have had to implement a layer on top of kociemba. My Kociemba solver basically will check if a cube is in \textit{standard} form, and if not, it will find the sequence of full cube rotations that bring us to an equivalent configuration in \textit{standard} form, pass that to equivalent \textit{standard} cube to kociemba, and stich the cube rotation to the kociemba solution (treating cube rotation as 0 cost everywhere in the code). That way, we seamlessly get to an answer that makes sense from kociemba, without the user having to worry about whether or not the cube they are passing as input is in \textit{standard} form or not! Running the same example as above via my KociembaSolver, we get the expected result:

\afblue
\paragraph{}{\textbf{Kociemba bug -- 3x3x3 massage to fix the broken interface}}
\begin{python}
#####################################################################
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
from rubiks.solvers.kociembasolver import KociembaSolver
from_kociemba = KociembaSolver.from_kociemba
to_kociemba = KociembaSolver.to_kociemba
#####################################################################
puzzle_type = Puzzle.rubiks_cube
n=3
init_from_random_goal=False
cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
solver_type = Solver.kociemba
solver = Solver.factory(**globals())
print(solver.solve(cube))
###################################################################
\end{python}
\black

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/KociembaBug333Fix1}
%
\includegraphics[scale=0.5]{./Figures/KociembaBug333Fix2}
%\decoRule
\caption[Kociemba Bug Fix]{3x3x3 Fixed}
\label{fig:KociembaBug333Fix}
\end{figure}

>>>>

\afblue
\paragraph{}{\textbf{Kociemba bug -- 2x2x2 \textbf{RC} -- example of good interface design}}
\begin{python}
#####################################################################
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
from rubiks.solvers.kociembasolver import KociembaSolver
from_kociemba = KociembaSolver.from_kociemba
to_kociemba = KociembaSolver.to_kociemba
#####################################################################
puzzle_type = Puzzle.rubiks_cube
n=2
init_from_random_goal=False
cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
solver_type = Solver.kociemba
solver = Solver.factory(**globals())
print(solver.solve(cube))
#####################################################################
\end{python}
\black

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{./Figures/KociembaBug222}
%\decoRule
\caption[Kociemba Bug]{2x2x2 \textbf{RC} -- hkociemba finds path to closest of the 24 equivalent goals}
\label{fig:KociembaBug222}
\end{figure}
As we can see, hkociemba happily returned a 1-move solution R to green F, white U, instead of the costlier solution it would have taken to get to \textit{standard} form red F white U (clearly at a cost of 3 with solution RUD').
\\
\\
In the 3x3x3 case, things are not quite that simple. There is no way to reproduce full cube rotation from faces rotations alone, since the center cubies will never move via the latter, but do via the former. If we run the equivalent of the above problem with the (3x3x3) kociemba library, we silently get to a false solution as is apparent from printing the resulting cube.
\afblue
\paragraph{}{\textbf{Kociemba bug -- 3x3x3 \textbf{RC} -- example of \textbf{bad} interface design}}
\begin{python}
#####################################################################
from kociemba import solve
#####################################################################
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.kociembasolver import KociembaSolver
from_kociemba = KociembaSolver.from_kociemba
to_kociemba = KociembaSolver.to_kociemba
#####################################################################
puzzle_type = Puzzle.rubiks_cube
n=3
init_from_random_goal=False
cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
print(cube)
cube_string = to_kociemba(cube)
solution = solve(cube_string)
print('Kociemba Solution:', solution)
print(cube.apply_moves(from_kociemba(solution)))
###################################################################
\end{python}
\black

\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{./Figures/KociembaBug333}
%\decoRule
\caption[Kociemba Bug]{3x3x3 \textbf{RC} -- kociemba does not check center cubies and returns wrong solution}
\label{fig:KociembaBug333}
\end{figure}
In order to counteract that, and make the comparisons between Kociemba 3x3x3 and other solvers and methods fairer, I have had to implement a layer on top of kociemba. My Kociemba solver basically will check if a cube is in \textit{standard} form, and if not, it will find the sequence of full cube rotations that bring us to an equivalent configuration in \textit{standard} form, pass that to equivalent \textit{standard} cube to kociemba, and stich the cube rotation to the kociemba solution (treating cube rotation as 0 cost everywhere in the code). That way, we seamlessly get to an answer that makes sense from kociemba, without the user having to worry about whether or not the cube they are passing as input is in \textit{standard} form or not! Running the same example as above via my KociembaSolver, we get the expected result, composed of 3 cube rotations to bring it to \textit{standard} form, followed by one face rotation to solve it:

\afblue
\paragraph{}{\textbf{Kociemba bug -- 3x3x3 \textbf{RC} --  massage to fix the broken interface}}
\begin{python}
#####################################################################
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
from rubiks.solvers.kociembasolver import KociembaSolver
from_kociemba = KociembaSolver.from_kociemba
to_kociemba = KociembaSolver.to_kociemba
#####################################################################
puzzle_type = Puzzle.rubiks_cube
n=3
init_from_random_goal=False
cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
solver_type = Solver.kociemba
solver = Solver.factory(**globals())
print(solver.solve(cube))
###################################################################
\end{python}
\black

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/KociembaBug333Fix1}
%\decoRule
\label{fig:KociembaBug333Fix1}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/KociembaBug333Fix2}
%\decoRule
\caption[Kociemba Bug Fix]{3x3x3 \textbf{RC} -- Fixed kociemba interface}
\label{fig:KociembaBug333Fix2}
\end{figure}
